Automatic Matching of Bullet Lands

### 2017-06-22 imaging-paper.Rnw
text size increased

im1 <- "images/Hamby (2009) Barrel/bullets/Br1 Bullet 1-5.x3p"
im2 <- "images/Hamby (2009) Barrel/bullets/Br1 Bullet 2-1.x3p"

@@

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\textwidth}\centering
\caption{Loess smooth of signatures  at a height of $x = 100\mu m$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
lofX$bulletname <- scrubPath(lofX$bullet)
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bulletname) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bulletname~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{Using a rolling median peaks and valleys are identified for each signature. Peaks and valleys on the signature correspond to striation marks on the bullet's surface. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  #lof <- processBullets(paths = images[c(5,7)], x = 100)

@@

  peaks$bulletname <- scrubPath(peaks$bullet)
  lofX$bulletname <- scrubPath(lofX$bullet)
  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), 
               data= peaks, alpha=0.7) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(bulletname~.) +
    geom_line(aes(x=y, y=l30, group=bulletname), data=lofX) +
  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) 
@
}
\end{subfigure}    
\begin{subfigure}[t]{\textwidth}\centering
    \caption{Rectangles in the back identify a striation mark on one of the bullets.  Matching striation marks are indicated by color filled rectangles and marked by an `o'. Mismatches are filled in grey and  marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.65\\textwidth', warning=FALSE>>=
peaks1$lines$bullet <- b12[1]
peaks2$lines$bullet <- b12[2]

@@

ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
  ylim(c(-6,6.5)) + theme_bw() +
  geom_line(data=lofX, aes(x=y, y=l30, group=bulletname, linetype=bulletname)) +
  scale_linetype_discrete("") +
  scale_colour_manual("", values=cols) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
  xlab("y")
@
    }
\end{subfigure}
\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between signatures (c).}
\end{figure}
%\afterpage{\clearpage}
Figure~\ref{fig:match} gives an overview of the automated matching routine: 
We first identify a stable region for each bullet land and extract the signature at the lowest height in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see Supplement Section~\ref{supp:bulletbottom} for a more detailed discussion).

@@

\subsection{Horizontal alignment}
Signatures of each of the two lands, 1-5 and 2-1,  in Figure~\ref{fig:manualmatch-rgl} are shown in Figure~\ref{fig:cross100} extracted at a height of $x = 100\mu m$. Striation marks show up in these representations as peaks and valleys.  The individual characteristics are prominent and, again, suggest a match between the lands. A horizontal shift of one of the signatures (result shown in Fig~\ref{fig:crosscutX}) emphasizes the strong similarities between signatures. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
lof$bulletname <- scrubPath(lof$bullet)
qplot(y, resid, group=bulletname, colour=bulletname, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned signatures.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
lofY$bulletname <- scrubPath(lofY$bullet)
qplot(y+min(lof$y), resid, group=bulletname, colour=bulletname, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 taken at  heights of $x = 100\mu m$. A horizontal shift of the values of bullet~1-5 to the right shows the similarity of the striation marks.}
\end{figure}
%
For this alignment we use the cross-correlation function to find a maximal amount of agreement between the signatures \citep{bachrach:2002, chu:2010, vorburger:2011, thompson:2013}.
This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

\begin{figure}[hbtp]
  \centering
<<ccf, echo=FALSE, fig.width = 8, fig.height = 3, out.width = '.65\\textwidth'>>=
lag <- ccf$lag[which.max(ccf$acf)]
qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75))
@
\caption{\label{fig:ccf}Cross-correlation function between the two signatures shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. The correlation is maximized at a lag of -2, indicating the largest amount of agreement between the signatures. Figure~\ref{fig:crosscutX} shows the slight change resulting from the lag-shifted signatures.}
\end{figure}

@@

lop <- lapply(reslist, function(res) {
ggplot() +
  theme_bw() + 
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5.5, data=res$lines, alpha=0.2) +
  geom_line(aes(x = y, y = l30, linetype = bullet),   data = res$bullets) +
  scale_linetype_discrete("") +
#  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylim(c(-6,6.5)) + ylab("") + xlab("") + 
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match))
})

@@

\begin{figure}[hbtp]
  \centering
<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.5\\textwidth', warning=FALSE>>=
br111.groove <- get_grooves(br111_bullet)
br111.loess <- fit_loess(br111_bullet, br111.groove)

@@

ggplot() + theme_bw() +
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
            data=peaks,  alpha=0.2) +
  geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks, alpha=.7) +
  scale_colour_brewer(palette="Set2") + 
  scale_fill_brewer(palette="Set2") +
  theme(legend.position="none") + 
  facet_grid(smoothfactor~., labeller="label_both") +
  geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
            colour="grey50") +
  geom_line(aes(x=y, y=smoothed), data=smooths) +
  ylab(expression(paste("Signature values (in ",mu,"m)", sep=""))) 
@
\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 45, respectively. Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 45 might over-smooth and cause the peaks/valleys to either end disappear or shift  horizontally from their original position in the signature.}
\end{figure}

@@

qplot(cmsdist, geom="bar") + theme_bw() + xlab("Number of CMS")
@
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Overlaid signatures of C-3 and the land with the top matching CMS.}
<<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
res <- reslist[[which.max(cmsdist)]]  
#res <- reslist[[which.max(cmsdist[-which.max(cmsdist)])]]  # number 2
res$bullets$bullet <- scrubPath(res$bullets$bullet)

print(ggplot() +
  theme_bw() + 
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
  scale_linetype_discrete("") + ylab("") +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
  ylim(c(-6,6)) +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))

@
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Top 2 match with C-3 based on CMS.}
<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
res <- reslist[[which(cmsdist==14)]]  # number 2
res$bullets$bullet <- scrubPath(res$bullets$bullet)

print(ggplot() +
  theme_bw() + 
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
  scale_linetype_discrete("") + ylab("") +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
  ylim(c(-6,6)) +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))

@@

dm <- melt(densities, measure.var=c("KM", "KNM"))
dm$var <- factor(dm$var, levels=features)
levels(dm$var)[3:4] <- c("#matches", "#non-matches")
levels(dm$variable) <- c("Known matches (KM)", "Known non-matches (KNM)")
dm$varLabel <- dm$var
levels(dm$varLabel) <- c(
  "Consecutive Matching Striae (CMS)", 
  "Consecutive Non-Matching Striae (CNMS)",
  "#matches",
  "#non-matches",
  "Average difference (D)",
  "Sum of peaks (S)",
  "Cross-correlation function (ccf)")

qplot(x, value, group=variable, geom="polygon", data=dm, fill=variable, 
      alpha=I(0.6), colour=I("grey20")) + 
  facet_wrap(~varLabel, ncol=4, scales="free") +
  theme_bw() + ylab("") + xlab("") +
  scale_fill_brewer("Matches", palette="Paired") +
  theme(legend.position=c(1,0), 
        legend.justification = c("right", "bottom"),
        legend.background = element_rect(colour="grey75"),
        axis.title.y=element_blank(),
        plot.margin = unit(c(0,0,0,0), unit="cm"))
@
\caption{\label{fig:densities}Overview of all the marginal densities for features described in section~\ref{sec:algorithm}. Shifts in the mode of the density functions between known matches and known non-matches indicate the variable's predictive power in distinguishing matches and non-matches. Predictive power is shown in more detail in Figure~\ref{fig:rocs}.}
\end{figure}

@@

set.seed(20140105)
aucs <- plyr::ldply(features, function(x) {
  subbstats <- subset(bstats, flagged == FALSE)
  pos.scores <- sample(subbstats[which(subbstats$match),x], 50000, replace=TRUE)
  neg.scores <- sample(subbstats[which(!subbstats$match),x], 50000, replace=TRUE)
  data.frame(x, upper=mean(pos.scores > neg.scores), 
             lower = mean(pos.scores < neg.scores))
})
aucs <- melt(aucs, measure.var=c("lower", "upper"))
names(aucs) <- c("variable", "type", "auc")
rocs <- merge(rocs, aucs, by=c("variable", "type"))
rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
levels(rocs$variable)[3:4] <- c("#non-matches", "#matches")
rocs$varLabel <- rocs$variable
levels(rocs$varLabel) <- c(
  "Sum of peaks (S)",
  "Cross-correlation function (ccf)",
  "#non-matches",
  "#matches",
  "Consecutive Non-Matching Striae (CNMS)",
  "Consecutive Matching Striae (CMS)", 
  "Average difference (D)"
  )

@@

ggplot(data=subset(rocs, auc > 0.5)) +
  geom_ribbon(aes(x=`FALSE`, ymax=`TRUE`, ymin=0), alpha=0.2) +
  geom_line(aes(x=`FALSE`,y=`TRUE`))+ theme_bw() + 
  xlab("False negative rate") + 
  ylab("False positive rate") + facet_wrap(facets=~varLabel, ncol=4) + 
  geom_label(x=1, y=0, aes(label=auc_lbl), data=eer, hjust=1, vjust=0) + 
  geom_abline(linetype=2) + 
  geom_point(aes(x=`FALSE`, y=`TRUE`), data=eer, size=2.5) +
  geom_label(x=1, y=0.125, aes(label=labels), data=eer, hjust=1, vjust=0)
@
\caption{\label{fig:rocs}ROC curves for all of the features described in section~\ref{sec:algorithm}. Variables are sorted according to their area under the curve (AUC). The equal error rate (EER) is marked by a point on the ROC curve. Except for the distance $D$ between signatures, all individual features derived from the surface measurements and the aligned striation marks are more predictive than the maximal CMS. }
\end{figure}

>>>>

im1 <- file.path(datadir, "Br1 Bullet 1-5.x3p")
im2 <- file.path(datadir, "Br1 Bullet 2-1.x3p")

@@

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\textwidth}\centering
\caption{Loess smooth of signatures  at a height of $x = 100\mu m$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
lofX$bulletname <- scrubPath(lofX$bullet)
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bulletname) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bulletname~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + #ylab("") + 
  theme(legend.position="none") +
    ylab(expression(paste("Residuals (in ",mu,m,")", sep=""))) + 
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
  theme(text = element_text(size=12))

@
    }
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{Using a rolling median peaks and valleys are identified for each signature. Peaks and valleys on the signature correspond to striation marks on the bullet's surface. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  #lof <- processBullets(paths = images[c(5,7)], x = 100)

@@

  peaks$bulletname <- scrubPath(peaks$bullet)
  lofX$bulletname <- scrubPath(lofX$bullet)
  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), 
               data= peaks, alpha=0.7) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(bulletname~.) +
    geom_line(aes(x=y, y=l30, group=bulletname), data=lofX) +
    ylab(expression(paste("Residuals (in ",mu,m,")", sep=""))) + 
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
  theme(text = element_text(size=12))
@
}
\end{subfigure}    
\begin{subfigure}[t]{\textwidth}\centering
    \caption{Rectangles in the back identify a striation mark on one of the bullets.  Matching striation marks are indicated by color filled rectangles and marked by an `o'. Mismatches are filled in grey and  marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.65\\textwidth', warning=FALSE>>=
peaks1$lines$bullet <- b12[1]
peaks2$lines$bullet <- b12[2]

@@

ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
  ylim(c(-6,6.5)) + theme_bw() +
  geom_line(data=lofX, aes(x=y, y=l30, group=bulletname, linetype=bulletname)) +
  scale_linetype_discrete("") +
  scale_colour_manual("", values=cols) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
  theme(text = element_text(size=12)) 
@
    }
\end{subfigure}
\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between signatures (c).}
\end{figure}
%\afterpage{\clearpage}
Figure~\ref{fig:match} gives an overview of the automated matching routine: 
We first identify a stable region for each bullet land and extract the signature at the lowest height in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see Supplement Section~\ref{supp:bulletbottom} for a more detailed discussion).

@@

\subsection{Horizontal alignment}
Signatures of each of the two lands, 1-5 and 2-1,  in Figure~\ref{fig:manualmatch-rgl} are shown in Figure~\ref{fig:cross100} extracted at a height of $x = 100\mu m$. Striation marks show up in these representations as peaks and valleys.  The individual characteristics are prominent and, again, suggest a match between the lands. A horizontal shift of one of the signatures (result shown in Fig~\ref{fig:crosscutX}) emphasizes the strong similarities between signatures. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
lof$bulletname <- scrubPath(lof$bullet)
qplot(y, resid, group=bulletname, colour=bulletname, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
  theme(text = element_text(size=20)) 
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned signatures.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
lofY$bulletname <- scrubPath(lofY$bullet)
qplot(y+min(lof$y), resid, group=bulletname, colour=bulletname, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
  theme(text = element_text(size=20)) 
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 taken at  heights of $x = 100\mu m$. A horizontal shift of the values of bullet~1-5 to the right shows the similarity of the striation marks.}
\end{figure}
%
For this alignment we use the cross-correlation function to find a maximal amount of agreement between the signatures \citep{bachrach:2002, chu:2010, vorburger:2011, thompson:2013}.
This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

\begin{figure}[hbtp]
  \centering
<<ccf, echo=FALSE, fig.width = 8, fig.height = 3, out.width = '.65\\textwidth'>>=
maxlag <- ccf$lag[which.max(ccf$acf)]
dframe <- data.frame(lag = ccf$lag, acf = ccf$acf)
qplot(data = dframe, x=lag, xend=lag, y = 0, yend= acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = maxlag, xend = maxlag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, maxlag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75)) +
  theme(text = element_text(size=20)) 
@
\caption{\label{fig:ccf}Cross-correlation function between the two signatures shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. The correlation is maximized at a lag of -2, indicating the largest amount of agreement between the signatures. Figure~\ref{fig:crosscutX} shows the slight change resulting from the lag-shifted signatures.}
\end{figure}

@@

lop <- lapply(reslist, function(res) {
ggplot() +
  theme_bw() + 
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5.5, data=res$lines, alpha=0.2) +
  geom_line(aes(x = y, y = l30, linetype = bullet),   data = res$bullets) +
  scale_linetype_discrete("") +
#  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylim(c(-6,6.5)) + ylab("") + xlab("") + 
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
#      ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
#  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
  theme(text = element_text(size=20)) 
})

@@

\begin{figure}[hbtp]
  \centering
<<smoothfac1, echo=FALSE, fig.width=10, fig.height=5.2, out.width='.75\\textwidth', warning=FALSE>>=
br111.groove <- get_grooves(br111_bullet)
br111.loess <- fit_loess(br111_bullet, br111.groove)

@@

ggplot() + theme_bw() +
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
            data=peaks,  alpha=0.2) +
  geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks, alpha=.7) +
  scale_colour_brewer(palette="Set2") + 
  scale_fill_brewer(palette="Set2") +
  theme(legend.position="none") + 
  facet_grid(smoothfactor~., labeller="label_both") +
  geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
            colour="grey50") +
  geom_line(aes(x=y, y=smoothed), data=smooths) +
  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
  theme(text = element_text(size=20)) 
@
\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 45, respectively. Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 45 might over-smooth and cause the peaks/valleys to either end disappear or shift  horizontally from their original position in the signature.}
\end{figure}

@@

qplot(cmsdist, geom="bar") + theme_bw() + xlab("Number of CMS") + 
  theme(text = element_text(size=20)) 
@
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Overlaid signatures of C-3 and the land with the top matching CMS.}
<<top, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth', warning = FALSE>>=
res <- reslist[[which.max(cmsdist)]]  
#res <- reslist[[which.max(cmsdist[-which.max(cmsdist)])]]  # number 2
res$bullets$bullet <- scrubPath(res$bullets$bullet)

ggplot() +
  theme_bw() + 
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, data=res$lines, alpha=0.2, show.legend=FALSE) +
  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
  scale_linetype_discrete("") + ylab("") +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
  ylim(c(-6,7)) +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
    ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
  theme(text = element_text(size=20))
@
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Top 2 match with C-3 based on CMS.}
<<top2, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth', warning = FALSE>>=
res <- reslist[[which(cmsdist==14)]]  # number 2
res$bullets$bullet <- scrubPath(res$bullets$bullet)

ggplot() +
  theme_bw() + 
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, data=res$lines, alpha=0.2, show.legend=FALSE) +
  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
  scale_linetype_discrete("") + ylab("") +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
  ylim(c(-6,7)) +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
    ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
  theme(text = element_text(size=20))

@@

dm <- melt(densities, measure.var=c("KM", "KNM"))
dm$var <- factor(dm$var, levels=features)
levels(dm$var)[3:4] <- c("#matches", "#non-matches")
levels(dm$variable) <- c("Known matches (KM)", "Known non-matches\n (KNM)")
dm$varLabel <- dm$var
levels(dm$varLabel) <- c(
  "Consecutive Matching\n Striae (CMS)", 
  "Consecutive Non-Matching\n Striae (CNMS)",
  "#matches",
  "#non-matches",
  "Average difference (D)",
  "Sum of peaks (S)",
  "Cross-correlation\n function (ccf)")

qplot(x, value, group=variable, geom="polygon", data=dm, fill=variable, 
      alpha=I(0.6), colour=I("grey20")) + 
  facet_wrap(~varLabel, ncol=4, scales="free") +
  theme_bw() + ylab("") + xlab("") +
  scale_fill_brewer("Matches", palette="Paired") +
  theme(legend.position=c(1,0), 
        legend.justification = c("right", "bottom"),
        legend.background = element_rect(colour="grey75"),
        axis.title.y=element_blank(),
        plot.margin = unit(c(0,0,0,0), unit="cm")) +
  theme(text = element_text(size=20)) 
@
\caption{\label{fig:densities}Overview of all the marginal densities for features described in section~\ref{sec:algorithm}. Shifts in the mode of the density functions between known matches and known non-matches indicate the variable's predictive power in distinguishing matches and non-matches. Predictive power is shown in more detail in Figure~\ref{fig:rocs}.}
\end{figure}

@@

set.seed(20140105)
aucs <- plyr::ldply(features, function(x) {
  subbstats <- subset(bstats, flagged == FALSE)
  pos.scores <- sample(subbstats[which(subbstats$match),x], 50000, replace=TRUE)
  neg.scores <- sample(subbstats[which(!subbstats$match),x], 50000, replace=TRUE)
  data.frame(x, upper=mean(pos.scores > neg.scores), 
             lower = mean(pos.scores < neg.scores))
})
aucs <- melt(aucs, measure.var=c("lower", "upper"))
names(aucs) <- c("variable", "type", "auc")
rocs <- merge(rocs, aucs, by=c("variable", "type"))
rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
levels(rocs$variable)[3:4] <- c("#non-matches", "#matches")
rocs$varLabel <- rocs$variable
levels(rocs$varLabel) <- c(
  "Sum of peaks (S)",
  "Cross-correlation\n function (ccf)",
  "#non-matches",
  "#matches",
  "Consecutive Non-Matching\n Striae (CNMS)",
  "Consecutive Matching\n Striae (CMS)", 
  "Average difference (D)"
  )

@@

ggplot(data=subset(rocs, auc > 0.5)) +
  geom_ribbon(aes(x=`FALSE`, ymax=`TRUE`, ymin=0), alpha=0.2) +
  geom_line(aes(x=`FALSE`,y=`TRUE`))+ theme_bw() + 
  xlab("False negative rate") + 
  ylab("False positive rate") + facet_wrap(facets=~varLabel, ncol=4) + 
  geom_label(x=1, y=0, aes(label=auc_lbl), data=eer, hjust=1, vjust=0) + 
  geom_abline(linetype=2) + 
  geom_point(aes(x=`FALSE`, y=`TRUE`), data=eer, size=2.5) +
  geom_label(x=1, y=0.125, aes(label=labels), data=eer, hjust=1, vjust=0) +
  theme(text = element_text(size=20)) 
@
\caption{\label{fig:rocs}ROC curves for all of the features described in section~\ref{sec:algorithm}. Variables are sorted according to their area under the curve (AUC). The equal error rate (EER) is marked by a point on the ROC curve. Except for the distance $D$ between signatures, all individual features derived from the surface measurements and the aligned striation marks are more predictive than the maximal CMS. }
\end{figure}
